Refinements for open automata

### 2022-11-30 2021-Refinement/paper/refinementarticleV2.tex
version 2 du papier avec def revisee

We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation. This leads us to the following definition.
\begin{definition}[Preservation along identical transitions]\label{def:preservation} \\
Let \(\OAg[1]\) and \(\OAg[2]\) be two OAs.
A relation \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) between  \(\OAg[1]\) and \(\OAg[2]\)
  is preserved along identical transitions if  it satisfies the following:
% TODO: isn't it a2 = a1x instead of a1x = a1
% TODO: ANS:  we have a1 = a2 so whatever, it is the same, we can change if it is more ``natural''
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} \!=\! \OTx{1}{x}{1x}{1x} \!\in\! \fOT{s_1}}^{x \in X}\!,
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\!\fvars{t_{1x}} \uplus \fvars{t_2} \uplus \fvars{t_1} \vdash \hspace{6cm}\\  R\mpar{s_1, s_2} \wedge 	g_1 \wedge g_2 \wedge \alpha_1 = \alpha_2 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_2} \beta_{1j} = \beta_{2j} 
\implies \\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigvee_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_2 \land \bigwedge_\subbox{j \in J'_2 } \beta_{2j} = \beta_{1xj}\land g_{1 x}\land
 R\mpar{s'_{1 x}, s'_2}\psubst{\psi_{1 x} \uplus \psi_2}}\\
\end{multline*}
\end{definition}

We define now a simulation between OA that have the same set of holes. This relation thus focuses on the simulation of transitions and uses the previous definition to prevent the refined automaton from non-deterministically reaching states that introduce deadlocks in a scenario planned without deadlock in the specification.
\begin{definition}[Hole-equal simulation]
Consider two OA with identical set of holes:  \(\OAg[1]\) and \(\OAg[2]\) where  \(J_1 = J_2\), the relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $S_1$ to $S_2$ if the following conditions hold: 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
\everymath{\displaystyle}\begin{array}{l}
		\bigsymb{\forall} t_1 = \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{t_{2x} = \OTx{2}{x}{2x}{2x} \in \fOT{s_2}}^{x \in X}, \\[10pt]
		\quad \mpar{\forall x \in X, J'_{2x}= J'_1} \nwedge\\[2pt]
		 V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\
\hspace{5em} \left( R\mpar{s_1, s_2} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
			\alpha_1 = \alpha_{2x} \wedge \bigwedge_\subbox{j \in J'_{2x}} \beta_{1j} = \beta_{2xj} \nwedge\\[12pt]
			 g_{2x} \wedge R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}}
		\end{array}}\right)
	\end{array} 
%\\
%	\wedge \mpar{V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}}
\end{multline*}
\item[(3)] $R$ is  is preserved along identical transitions of \(A_1\) and \(A_2\).
\end{definition}

\TODO{Simulation: Il y a differentes facons de voir le relation de reffinement entre l'implem et la spec:
\begin{itemize}
\item implem as a conformance of spec:  implem deadlocks less often.
\item implem as a (nondeterminism) reduction of spec:  implem deadlocks less often. Can only used be used for complete specification.
\item implem as an extension of spec. C'est le complement du raffinement par reduction, i.e. Can be used to complete partial specification (implem can exhibit more behaviour than spec).
Our definition of simulation matches the definition given in \cite{EshuisF:2002} Section 3.2 $\implies$ Refinement par extension according to Thm 8 (Section 4)
\end{itemize}
}

@@

\section{Proof of Transitivity for Refinement}\label{sec:proof-transitivity}
\begin{quote}
If $\wrel{A_1}{A_2}{H}$ and $\wrel{A_2}{A_3}{H'}$, then $\wrel{A_1}{A_3}{H\cap H'}$.
\end{quote}
\proof 
If $\wrel{A_1}{A_2}{H}$ then there is $R_{12}$ a relation between states
of $A_1$ and of $A_2$;  If $\wrel{A_2}{A_3}{H'}$ then there is $R_{23}$ a relation between states of $A_2$ and of $A_3$. We build a relation between
 states of $A_1$ and of $A_3$ as follows:  for each pair of states $s_1$, $s_3$, for each state $s_2$ such that $R_{12}$ relates $s_1$ and $s_2$, and $R_{23}$ relates $s_2$ and $s_3$.
Let $R_{13}$ be the relation:\\
  \[R_{13}(s_1,s_3)=\bigvee_{s_2\in S_2}\mpar{R_{12}(s_1,s_2) \land R_{23}(s_2,s_3) } \]

We will show that $\wrel{A_1}{A_3}{H\cap H'}$ by exhibiting  $R_{13}$ as a hole-tracking simulation of $A_1$ by  $A_3$.

We have to prove that the relation $R_{13}$ satisfies the three conditions of the definition of a refinement of open automata.
\begin{enumerate}
\item Firstly, we have to $R_{13}$ satisfies initial configurations:
\[\sigma_{01} \uplus \sigma_{03} \vdash R_{13} \mpar{s_{01}, s_{03}}\]
By knowing that substitutions only have an effect on the variables of the open automaton they belong to, they also produce terms containing only variables of the open automaton they belong to. We have:
\begin{align*}
\mpar{\sigma_{01} \uplus \sigma_{02} \vdash R_{12}\mpar{s_{01}, s_{02}}} \wedge \mpar
{\sigma_{02} \uplus \sigma_{03} \vdash R_{23}\mpar{s_{02}, s_{03}}}&\implies\\
R_{12}\mpar{s_{01},s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02}} \wedge R_{23}\mpar{s_{02}, s_{03}}\psubst{\sigma_{02} \uplus \sigma_{03}}&\implies\\
R_{12}\mpar{s_{01},s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}} \wedge R_{23}\mpar{s_{02}, s_{03}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}}&\implies\\ 
\underbrace{R_{12}\mpar{s_{01},s_{02}}\wedge R_{23}\mpar{s_{02}, s_{03}}}_{\implies R_{13} \mpar{s_{01}, s_{03}}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03}} %&\implies\\
 \end{align*}
Since $\sigma_{02}$ has no effect on variables of $s_{01}$ and $s_{03}$ thus we get the expected result.

\item Secondly, we need to prove that for any open transition $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
there exists an indexed family of OTs originating from $s_{3}$: 
\begin{multline*}
V_1 \uplus  V_3 \uplus \fvars{t_1}
\vdash\\\hspace{2em}
\mpar{R_{13}\mpar{s_1, s_3} \wedge g_1 \implies 
\operatorname*{\bigsymb{\bigvee}}_{z \in Z}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3z}\wedge \bigwedge_\subbox{j \in J'_{3z} \cap (H\cap H')} \beta_{1j}=\beta_{3jz} \wedge \\g_{3z}\wedge 
{R_{12}\mpar{s'_1,s'_{3z}}}\psubst{\psi_1  \uplus \psi_{3z}}
\end{array}}}
\end{multline*}	
\medskip
 Consider $\mpar{s_{1}, s_{3}} \in R_{13}$ then there is a set of states $(s_{2p})^{p\in P}$ of $A_2$ relating  $s_{1}$ and $s_{3}$:
\[R_{13}(s_1,s_3)=\bigvee_{{p\in P}
}\mpar{R_{12}\mpar{s_1,s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}} \]

Let's consider any $s_{2p} \in (s_{2p})^{p\in P}$. We have on one side,  for any open transition $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
there exists an indexed family of OTs originating from $s_{2p}$: 
\begin{align*}
\mpar{\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{x \in X_p} 
\end{align*}
such that $\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$ and
\begin{multline*}
%\everymath{\displaystyle}
V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \nwedge\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} 
\end{multline*}		
Adding to both sides of the implication the predicate $R_{23}\mpar{s_{2p},s_3}$ we get: 
\begin{multline*}
V_1 \uplus V_2 \uplus V_3  \uplus \fvars{t_1} \vdash\\
R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3} \wedge g_1 \implies\hspace{5.5cm}\\ \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \land\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}} \land R_{23}\mpar{s_{2p},s_3}
		\end{array}} \qquad (*) 
\end{multline*}	
On the other side, according the relation between $A_2$ and $A_3$ we have for any open transition $t_{2px}$ in $T_2$ originating from $s_{2p}$:
\begin{align*}
\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}
\end{align*}
there exists an indexed family of OTs originating from $s_3$: 
\begin{align*}
\mpar{\OTx{3}{pxy}{3pxy}{3pxy} \in \fOT{s_3}}^{y \in Y} 
\end{align*}
such that $\forall y \in Y, J'_{2px} \cap H' = J'_{3pxy} \cap H'$ and
\begin{multline*}
%\everymath{\displaystyle}
V_2 \uplus V_3 \uplus \fvars{t_{2px}} \vdash\\
R_{23}\mpar{s_{2p}, s_{3}} \wedge g_{2px} \implies \hspace{7cm}\\ \operatorname*{\bigsymb{\bigvee}}_{y \in Y} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			 \alpha_{2px}=\alpha_{3pxy} \wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap H'} \beta_{2pxj}=\beta_{3pxyj} \nwedge \\
			g_{3pxy} \wedge R_{23}\mpar{s'_{2px}, s'_{3xy}}\psubst{\psi_{2px} \uplus \psi_{3pxy}}
		\end{array}}  \qquad (**)
\end{multline*}

From the  two previous cases: 
$\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$ and
$\forall y \in Y, J'_{2px} \cap H' = J'_{3pxy} \cap H'$ we conclude:
$\forall x \in X, \forall y \in Y,  J'_1 \cap (H \cap H') = J'_{3pxy} \cap (H\cap H')$ \\

In addition, by combining  formula $(**)$ and $(*)$, we get: 
\begin{multline*}
V_1 \uplus V_2 \uplus V_3  \uplus \fvars{t_1}\uplus \fvars{t_{2px}} \vdash\\
R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3} \wedge g_1 \implies \hspace{4.5cm}\\\operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj}  \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}} \\ 
\land
		\operatorname*{\bigsymb{\bigvee}}_{y \in Y} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			 \alpha_{2px}=\alpha_{3pxy} \wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap H'} \beta_{2pxj}=\beta_{3pxyj} \nwedge\\
			 g_{3pxy} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}\psubst{\psi_{2px} \uplus \psi_{3pxy}}
		\end{array}}	
		\end{array}} 
\end{multline*}		
With  the rearrangement of the formula, we get:
\begin{multline*}
V_1 \uplus V_2 \uplus V_3  \uplus \fvars{t_1}\uplus \fvars{t_{2px}}  \vdash\\
R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3} \wedge g_1 \implies 
\hspace{5.5cm}\\
\operatorname*{\bigsymb{\bigvee}}_{x \in X} 
\operatorname*{\bigsymb{\bigvee}}_{y \in Y}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3pxy}\wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap (H\cap H')} \beta_{1j}=\beta_{3jpxy} \wedge g_{3pxy}\wedge \\
R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}\psubst{\psi_{2px} \uplus \psi_{3pxy}}
\end{array}}
\end{multline*}

Because of the domain of the substitutions of the relations, the formula can be re-written:
\begin{multline*}
V_1 \uplus V_2 \uplus V_3  \uplus \fvars{t_1}\uplus \fvars{t_{2px}} \vdash\\
R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3} \wedge g_1 \implies 
\hspace{5.5cm}\\
\operatorname*{\bigsymb{\bigvee}}_{x \in X} 
\operatorname*{\bigsymb{\bigvee}}_{y \in Y}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3pxy}\wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap (H\cap H')} \beta_{1j}=\beta_{3jpxy} \wedge g_{3pxy}\wedge \\
\mpar{R_{12}\mpar{s'_1, s'_{2px}} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}}\psubst{\psi_1 \uplus \psi_{2px} \uplus \psi_{3pxy}}
\end{array}}
\end{multline*}
The formula is valid for all $s_{2p} \in (s_{2p})^{p\in P}$. To build $R_{13}$ we need to rely on the disjunction of all possible paths to relate $s_1$ and $s_3$, which leads to\\ ${\displaystyle R_{13}(s_1,s_3)=\bigvee_{{p\in P}}\mpar{R_{12}\mpar{s_1,s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}}}$.
\begin{multline*}
V_1 \uplus V_2 \uplus V_3  \uplus \fvars{t_1} \uplus  \biguplus_\subbox{p \in P} \fvars{t_{2px}} \vdash\\
\operatorname*{\bigsymb{\bigvee}}_{p\in P} \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3} \wedge g_1} \implies 
\hspace{5.5cm}\\
\operatorname*{\bigsymb{\bigvee}}_{p\in P}
\operatorname*{\bigsymb{\bigvee}}_{x \in X} 
\operatorname*{\bigsymb{\bigvee}}_{y \in Y}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3pxy}\wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap (H\cap H')} \beta_{1j}=\beta_{3jpxy} \wedge g_{3pxy}\wedge \\
\underbrace{R_{12}\mpar{s'_1, s'_{2px}} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}}_{\implies R_{13}\mpar{s'_1,s'_{3pxy}}}\psubst{\psi_1 \uplus \psi_{2px} \uplus \psi_{3pxy}}
\end{array}}
\end{multline*}
Indeed, since $\psi_{2px}$ has no effect on variables of $A_1$ and $A_3$ we have:
\begin{align*}
R_{12}\mpar{s'_1, s'_{2px}} \wedge R_{23}\mpar{s'_{2px}, s'_{3pxy}}\psubst{\psi_1 \uplus \psi_{2px} \uplus \psi_{3pxy}}  \\  
\implies R_{13}\mpar{s'_1,s'_{3pxy}}
\psubst{\psi_1 \uplus \psi_{2px} \uplus \psi_{3pxy}}  \\ 
\implies R_{13}\mpar{s'_1,s'_{3pxy}}\psubst{\psi_1  \uplus \psi_{3pxy}}
\end{align*}
This allows us to conclude that the family of open transitions of $A_3$ indexed over $p$, $x$,  and $y$ simulates the original open transition  with note that $\psi_{2px}$ has no effect on variables of $A_1$ and $A_3$:
\begin{multline*}
V_1  \uplus V_3  \uplus \fvars{t_1} \vdash\\
R_{13}(s_1,s_3)  \wedge g_1 \implies 
\hspace{5.5cm}\\
\operatorname*{\bigsymb{\bigvee}}_{p\in P}
\operatorname*{\bigsymb{\bigvee}}_{x \in X} 
\operatorname*{\bigsymb{\bigvee}}_{y \in Y}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3pxy}\wedge \bigwedge_\subbox{j \in J'_{3pxy} \cap (H\cap H')} \beta_{1j}=\beta_{3jpxy} \\ \wedge g_{3pxy} 
{\wedge R_{13}\mpar{s'_1,s'_{3pxy}}}\psubst{\psi_1  \uplus \psi_{3pxy}}
\end{array}}
\end{multline*}
Overall, we have a family of open transitions $t^{p\in P,x\in X,y\in Y}_{pxy} \subseteq T_3$ that should simulate $t_1$. All  combinations of elements in $P$, $X$, and $Y$ provide a set $Z$ of  open transitions.  This allows us to get the desired conclusion, that there is a set of  open transitions indexed over $Z$:
%This provides the desired conclusion:\\
\begin{multline*}
V_1 \uplus  V_3 \uplus \fvars{t_1}
\vdash\\\hspace{2em}
\mpar{R_{13}\mpar{s_1, s_3} \wedge g_1 \implies 
\operatorname*{\bigsymb{\bigvee}}_{z \in Z}
\mpar{\everymath{\displaystyle}
\begin{array}{l}
\alpha_1=\alpha_{3z}\wedge \bigwedge_\subbox{j \in J'_{3z} \cap (H\cap H')} \beta_{1j}=\beta_{3jz} \wedge \\g_{3z}\wedge 
{R_{13}\mpar{s'_1,s'_{3z}}}\psubst{\psi_1  \uplus \psi_{3z}}
\end{array}}}
\end{multline*}

\item  Finally, we have to prove the following holds:
\begin{multline*}
\bigforall \mpar{s_1, s_3} \in S_1 \times S_3, \bigsymb{\forall} \mpar{t_1, t_3} = \mpar{\OTx{1}{}{1}{1}, \OTx{3}{}{3}{3}} \in \fOT{s_1} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1x} = \OTx{1}{x}{1x}{1x} \in \fOT{s_1}}^{x \in X},
 V_1 \uplus V_3\uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_1} \uplus \fvars{t_3} \vdash \hspace{6cm}\\  R_{13}\mpar{s_1, s_3} \wedge 	g_1 \wedge g_3 \wedge \alpha_1 = \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_3} \beta_{1j} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigvee_{x\in X}}
\mpar{
\alpha_1=\alpha_{1 x}  \land J'_{1x}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{1xj}\land g_{1 x}\land
 R_{13}\mpar{s'_{1 x}, s'_3}\psubst{\psi_{1 x} \uplus \psi_3}}
\end{multline*}

 Consider $\mpar{s_{1}, s_{3}} \in R_{13}$ then there is a set of states $(s_{2p})^{p\in P}$ of $A_2$ relating  $s_{1}$ and $s_{3}$:
\[R_{13}(s_1,s_3)=\bigvee_{{p\in P}
}\mpar{R_{12}\mpar{s_1,s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}} \]

Consider $(t_1,t_3)\in\fOT{s_1} \times \fOT{s_3}$

\newpage
\TODO{[RABEA] proof}
Let's consider any $s_{2p} \in (s_{2p})^{p\in P}$. Consider  $t_1$ in $T_1$  originating from $s_1$:
\begin{align*}
\OTx{1}{}{1}{1} \in \fOT{s_1}
\end{align*}
According to the refinement relation between  $A_1$ and $A_2$ there exists an indexed family of OTs originating from $s_{2p}$: 
\begin{align*}
\mpar{\OTx{2p}{x}{2px}{2px} \in \fOT{s_{2p}}}^{x \in X_p} 
\end{align*}
such that $\forall x \in X, J'_{2px} \cap H = J'_1 \cap H$ and
\begin{multline}\label{lab1}\tag{$*$}
%\everymath{\displaystyle}
V_1 \uplus V_2 \uplus \fvars{t_1} \vdash\\ \mpar{R_{12}\mpar{s_1, s_{2p}} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X_p} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px} \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \nwedge\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}}} 
\end{multline}	
On the other hand, according to the refinement relation between   $A_2$ and $A_3$, we know condition 3 is satisfied, $\bigforall \mpar{s_{2p}, s_3} \in S_2 \times S_3$ and $\bigsymb{\forall} \mpar{t_{2p}, t_3} \in \fOT{s_{2p}} \times \fOT{s_3}$ and in particular for  the set $(t_{2px})^{x\in X_p}$ so we  have the following that we denote ($**$):
\begin{multline*}
\bigsymb{\forall} \mpar{t_{2px}, t_3} = \mpar{\OTx{2px}{}{2px}{2px}, \OTx{3}{}{3}{3}} \in \fOT{s_{2p}} \times \fOT{s_3} \\ 
\bigsymb{\bigexists} 
\mpar{t_{2pxy} = \OTx{2px}{y}{2pxy}{2pxy} \in \fOT{s_{2p}}}^{y \in Y_X},
V_2 \uplus V_3\uplus \biguplus_{y\in Y}\fvars{t_{2pxy}} \uplus \fvars{t_3} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{23}\mpar{s_{2p}, s_3} \wedge 	g_{2px}\wedge g_3 \wedge \alpha_{2px}= \alpha_3 
%\wedge \bigwedge_\subbox{j \in J'_2 \cap J'_3} \beta_{2pj} = \beta_{3j} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_2}. \\
\bigsymb{\bigvee_{y\in Y_X}}
\mpar{
\alpha_{2px}=\alpha_{2p xy}  \land J'_{2pxy}= J'_3 \land \bigwedge_\subbox{j \in J'_3 } \beta_{3j} = \beta_{2pxyj}\land g_{2p xy}\land
 R_{23}\mpar{s'_{2p xy}, s'_3}\psubst{\psi_{2p xy} \uplus \psi_3}}\\
\end{multline*}
combining $(*)$ and $(**)$ with the definition of the set $s_{2p}$, we obtain (variables on the left  of the $\vdash$ omitted for readability)
\begin{align*}
 R_{13}\mpar{s_1, s_3} &\wedge 	g_1 \wedge g_3 \wedge \alpha_1 = \alpha_3 \implies \\
&(\bigvee_{{p\in P}}\mpar{R_{12}\mpar{s_1,s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}}) \wedge 	g_1 \wedge g_3 \wedge \alpha_1 = \alpha_3  \implies \\
& (\bigvee_{{p\in P}}{\bigsymb{\bigvee}}_{x \in X_p} \left(\mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_1 = \alpha_{2px}= \alpha_3 \wedge \bigwedge_\subbox{j \in J'_{2px} \cap H} \beta_{1j} = \beta_{2pxj} \nwedge\\
			 g_{2px} \wedge R_{12}\mpar{s'_1, s'_{2px}}\psubst{\psi_1 \uplus \psi_{2px}}
		\end{array}} \right) 
	\wedge R_{23}\mpar{s_{2p},s_3}  \wedge g_3  \implies \\
\end{align*}

Again according to the refinement  relation between  $A_1$ and $A_2$ and condition 3  the following is satisfied $\bigforall \mpar{s_1, s_{2p}} \in S_1 \times S_2$ , in particular over the family of transitions in $(t_{2pxy})^{y\in Y_X}$:
\begin{multline*}
\bigsymb{\forall} \mpar{t_1, t_{2pxy}} = \mpar{\OTx{1}{}{1}{1}, \OTx{2pxy}{}{2pxy}{2pxy}} \in \fOT{s_1} \times \fOT{s_{2p}} \\ 
\bigsymb{\bigexists} 
\mpar{t_{1z} = \OTx{1}{z}{1z}{1z} \in \fOT{s_1}}^{z \in Z},
 V_1 \uplus V_2 \uplus \biguplus_{x\in X}\fvars{t_{1x}} \uplus \fvars{t_1} \uplus \fvars{t_{2p}} \vdash \hspace{6cm}\\  R_{12}\mpar{s_1, s_{2p}} \wedge 	g_1 \wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy} 
%\wedge \bigwedge_\subbox{j \in J'_1 \cap J'_{2p}} \beta_{1j} = \beta_{2pxyj} 
\implies \hspace{6cm}\\
%\exists
%\OTx{1}{a}{3}{3}\in \fOT{s_1}. \\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}}
\end{multline*}

This is verified for all $y \in  Y_X$.
\begin{multline*}
\bigsymb{\bigwedge_{y \in  Y_X}}
\Bigg[R_{12}\mpar{s_1, s_{2p}} \wedge 	g_1 \wedge g_{2pxy} \wedge \alpha_1 = \alpha_{2pxy}
\implies \hspace{6cm}\\
\bigsymb{\bigwedge_{z\in Z}}
\mpar{
\alpha_1=\alpha_{1 z}  \land J'_{1z}= J'_{2pxy} \land \bigwedge_\subbox{j \in J'_{2pxy} } \beta_{2pxyj} = \beta_{1zj}\land g_{1 z}\land
 R_{12}\mpar{s'_{1 z}, s'_{2pxy}}\psubst{\psi_{1 z} \uplus \psi_{2pxy}}}\Bigg]
\end{multline*}

>>>>

We thus define the following property that can be understood as follows. Take a specification automaton and its implementation. Consider two related states for the refinement relation. Suppose an identical transition exist in both automata (with same guards and same label), then either the reached states $s_1'$ and $s_2'$ are related by refinement, or there exists another transition of the specification that is also identical and leads to a state $s_1''$ that is related to $s_2'$ in the refinement relation.
Finally, we should take into account that 1) instead of equality, we should check if there is a state such that both transitions are feasible (the conjunction of guards is satisfiable), and that  2) instead of finding another transition of the specification we might find a family of transitions that cover all the cases of the transition of the implementation.

We define now a simulation between OA that have the same set of holes. This relation thus focuses on the simulation of transitions and uses the previous definition to prevent the refined automaton from non-deterministically reaching states that introduce deadlocks in a scenario planned without deadlock in the specification.
\TODO{Changement de la definition de simulation pour selon la simulation de Olga}

\begin{definition}[Hole-equal simulation]
Consider two OA with identical set of holes:  \(\OAg[1]\) and \(\OAg[2]\) where  \(J_1 = J_2\), the relation on configurations \(R: S_1 \times S_2 \to \rformulas[V_1 \uplus V_2]\) is a hole-\(\symb{equal}\) simulation from $S_1$ to $S_2$ if the following conditions hold: 
\item[(1)] \(\sigma_{01} \uplus \sigma_{02} \vdash R\mpar{s_{01}, s_{02}}\)
\item[(2)] \(\forall \mpar{s_1, s_2} \in S_1 \times S_2,\)\vspace{-8pt}
\noindent\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, \bigsymb{\forall} \mpar{t_1, t_2} = \mpar{\OTx{1}{}{1}{1}, \OTx{2}{}{2}{2}} \in \fOT{s_1} \times \fOT{s_2} \\ 
\everymath{\displaystyle}\begin{array}{l}
V_1 \uplus V_2 \uplus \fvars{t_1} \uplus \fvars{t_2}  \vdash R\mpar{s_1, s_2} \wedge g_2 \implies\\
V_1 \uplus V_2 \uplus \fvars{t_{1x}} \uplus \fvars{t_2}\vdash  \bigsymb{\bigexists} 
\mpar{t_{1x} \!=\! \OTx{1}{x}{1x}{1x}}^{x \in X} \operatorname*{\bigsymb{\bigvee}}_{x \in X}
\mpar{\begin{array}{l}
			\alpha_{1x} = \alpha_{2} \wedge \bigwedge_\subbox{j \in J'_{2}} \beta_{1xj} = \beta_{2j} \nwedge\\[12pt]
			 g_{1x} \wedge R\mpar{s'_{1x}, s'_{2}}\psubst{\psi_{1x} \uplus \psi_{2}}
		\end{array}} 
\end{array} 
\end{multline*}
\item [(3)] The deadlock reduction:
\begin{multline*}
\forall \mpar{s_1, s_2} \in S_1 \times S_2, V_1 \uplus V_2  \fvars{t_2} \vdash R\mpar{s_1, s_2} \wedge  \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \implies \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}  
\end{multline*}

\end{definition}

\TODO{To discuss this definition}

@@

\section{Proof of Transitivity for Refinement}\label{sec:proof-transitivity}
